---
title: "Homework 3 new"
author: "Kiran Kui"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
library(rnoaa)
library(hexbin)
library(patchwork)
library(packrat)
library(ggridges)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

```{r}
#Loading the Instacart data
library(p8105.datasets)
data("instacart")

instacart

```

his dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order.

Variables include: 

* order id
* product id
* add to cart order
* reordered items
* user id
* evaluation set
* order number
* day of week that order was made
* hour of day that order was made
* days since last order
* product name 
* aisle id
* department id
* aisle 
* department

One of the key variables is the product name purchased at instacart: for instance, "shelled pistachios", "spring water". 

Another key variable is the aisle that the product was purchased: for instance, yoghurt, fresh vegetables, etc. 

In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

I created a table summarizing the number of items ordered from aisle. In total, there are 134 aisles. 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

I created a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

The three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```
I created a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

# Problem 2

I read in, tidied and wrangled the dataset accel_data.csv

```{r}
# Load, tidy and wrangle the data
accel_df <- read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>% 
  drop_na(day_id) %>% 
  pivot_longer( 
     cols = activity_1:activity_1440,
     names_to = "activity_minute",
     names_prefix = "activity_",
     values_to = "activity_amount") %>% 
  mutate(wkday_wkend = case_when(
        day == "Saturday" ~ "weekend",
        day == "Sunday" ~ "weekend",
        day == "Monday" ~ "weekday",
        day == "Tuesday" ~ "weekday",
        day == "Wednesday" ~ "weekday",
        day == "Thursday" ~ "weekday",
        day == "Friday" ~ "weekday",
        TRUE ~ ""),
        week = as.integer(week),
        day_id = as.integer(day_id), 
        activity_minute = as.numeric(activity_minute)) %>% 
  mutate(day=fct_relevel(day, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))

```

There are `r ncol(accel_df)` rows(observations) and `r nrow(accel_df)` columns variables in this dataset. This dataset describes the activity of a 63-year-old male with a BMI of 25, diagnosed with congestive heart failure at the Advanced Cardiac Care Center of Columbia University Medical Center. 

Each observation(row) aligns with a single activity of this individual. 

There are 6 variables `r colnames(accel_df)`corresponding to week, day ID, day of the week, activity minute, activity amount, and whether the activity was conducted on a weekday or a weekend, respectively.  

```{r}
# New table showing aggregation across minutes
accel_df_2 = accel_df %>% 
group_by(week, day) %>% 
  summarize(total_activity = sum(activity_amount)) %>% 
  pivot_wider(names_from = day,
              values_from = total_activity)
```

I aggregated across minutes to create a total activity variable for each day, and create a table showing these totals called accel_df_2. 

Some trends are apparent. My results show that physical activity was highest on Fridays with and lowest on Saturdays. 
